---
title: "Abby_Code"
author: "Abby Robinson"
date: "7/12/2022"
output: html_document
---

Load Libraries
```{r}
library(curl)
library(survival)
library(survminer)
library(ggplot2)
```

Load caterpillar development data using the "curl" command 
```{r}
c <- curl("https://raw.githubusercontent.com/butterfliesrcool/Limenitis-Host-Plant-Gene-Expression/main/cat.dev.data1.csv")
dev <- read.csv(c, header = TRUE, sep = ",", stringsAsFactors = FALSE)
dev
length(dev$caterpillar.length)
```

Remove NA data 
```{r}
dev <- na.omit(dev)
length(dev$caterpillar.length)
```

Summary Stats 
Code Source: https://r-coder.com/aggregate-r/
```{r}
only.cats <- subset(dev, stage == "caterpillar", drop = FALSE)

sum.stats <- aggregate(x= only.cats$caterpillar.length, by= list(only.cats$date, only.cats$host.plant), FUN=mean)
colnames(sum.stats)[1] <- "Date"
colnames(sum.stats)[2] <- "Host-Plant"
colnames(sum.stats)[3] <- "Caterpillar-Length(mm)"
sum.stats
```
```{r}
day1 <- subset(only.cats, date == "15-Jul", drop = FALSE)

day1.aspen <- subset(day1, host.plant == "aspen", drop = FALSE)
day1.aspen
length(day1.aspen$caterpillar.length)
```

```{r}
aspen <- subset(sum.stats, sum.stats[2] == "aspen", drop = FALSE)
aspen
aspen$`Caterpillar-Length(mm)`
```

```{r}
cherry <- subset(sum.stats, sum.stats[2] == "cherry", drop = FALSE)
cherry
cherry$`Caterpillar-Length(mm)`
```

```{r}
aspen[1]
```

```{r}
df <- data.frame(
  date = aspen[1],
  aspen = aspen$`Caterpillar-Length(mm)`,
  cherry = cherry$`Caterpillar-Length(mm)`
)

growth <- t(cbind(df$aspen, df$cherry))

barplot(growth, beside=T, names.arg = df$date, col=c("black","grey"))
```
**Survival Analysis**
Within the package "Survival" we can use the function "Surv" to create a Survival object. We can then use the function "survfit" to generate the rate of pupation 
Code Source: https://www.emilyzabor.com/tutorials/survival_analysis_in_r_tutorial.html
```{r}
pup.rate <- survfit(Surv(dev.days, status.) ~ host.plant, data = dev)
summary(pup.rate)
```


Cox proportional-hazards regression model 
```{r}
coxph(Surv(dev.days, status.) ~ host.plant, data = dev)
```

```{r}
ggsurvplot(
    fit = survfit(Surv(dev.days, status.) ~ host.plant, data = dev), 
    xlab = "Days to Pupation", 
    ylab = "Rate of Pupation",
    fun = "event")
```


```{r}
### non-ggplot plot (lots of easy resources online to make these pretty

cats <- survfit(Surv(dev.days, status.) ~ host.plant, data = dev)
plot(cats,     
    xlab = "days since hatching", 
    ylab = "rate of pupation",
    fun = "event",
    col = c("darkolivegreen", "darkorchid4" ), lwd=3,
    )

legend(0, 1.0, legend=c("Aspen", "Black Cherry"),
       col=c("darkolivegreen", "darkorchid4"), lty=1:1, cex=1.5, title = "Host Plants")

```

R Module for making pretty figures in ggplot: https://www.cedricscherer.com/2019/08/05/a-ggplot2-tutorial-for-beautiful-plotting-in-r/

things we can do next: 
1. Make the ggplot prettier
2. Write interpretation of the survival analysis 
3. Add to summary stats: calculate average days to pupation for caterpillars reared on aspen and cherry 4. Add instars to the dataset, pull out all 4th/5th instar caterpillars or pre-pupa and see if we see significant differences in length of pre-pupa caterpillars between cherry and aspen 
5. Make cool figures showing the differences in average caterpillar length between the two host plants (might be more informative to show only the length of late instar caterpillars)
6. Try to calculate caterpillar mortality by comparing the original caterpillar counts (in the google drive) to the number of pupa/adults we have 




