```{r}
library(curl)
library(survival)
library(survminer)
library(ggplot2)
```
Load caterpillar mortality data using the "curl" command (only for mortality plot)
```{r}
c <- curl("https://raw.githubusercontent.com/butterfliesrcool/Limenitis-Host-Plant-Gene-Expression/main/mort.data.for.survival.plot")
dev <- read.csv(c, header = TRUE, sep = ",", stringsAsFactors = FALSE)
dev
length(dev$caterpillar.length) 
```
```{r}
mortality <- survfit(Surv(dev.days, mort.status) ~ host.plant, data = dev)
summary(mortality)
```
Cox proportional-hazards regression model 
```{r}
coxph(Surv(dev.days, mort.status) ~ host.plant, data = dev)
```
```{r}
### non-ggplot plot (lots of easy resources online to make these pretty

mortality <- survfit(Surv(dev.days, mort.status) ~ host.plant, data = dev)
plot(mortality,     
    xlab = "days since hatching", 
    ylab = "rate of caterpillar death",
    col = c("darkolivegreen", "darkorchid4" ), lwd=3,
    )

legend(0, 1.0, legend=c("Aspen", "Black Cherry"),
       col=c("darkolivegreen", "darkorchid4"), lty=1:1, cex=1.5, title = "Host Plants")

segments(35, .372, 40, .372, col = c("darkolivegreen"), lwd=3)
segments(35, .104, 40, .104, col = c("darkorchid4"), lwd=3)
```
Load caterpillar development data using the "curl" command (for pupation rate analysis)
```{r}
c <- curl("https://raw.githubusercontent.com/butterfliesrcool/Limenitis-Host-Plant-Gene-Expression/main/cat.dev.mort.data.csv")
dev <- read.csv(c, header = TRUE, sep = ",", stringsAsFactors = FALSE)
dev
length(dev$caterpillar.length) 
```
**Survival Analysis**
Within the package "Survival" we can use the function "Surv" to create a Survival object. We can then use the function "survfit" to generate the rate of pupation 
Code Source: https://www.emilyzabor.com/tutorials/survival_analysis_in_r_tutorial.html

```{r}
pup.dev <- subset(dev, stage %in% c("caterpillar", "pupa", "adult"), drop = FALSE)
pup.dev
```

```{r}
pup.rate <- survfit(Surv(dev.days, pupa.status) ~ host.plant, data = pup.dev)
summary(pup.rate)
```
Cox proportional-hazards regression model 
```{r}
coxph(Surv(dev.days, pupa.status) ~ host.plant, data = pup.dev)
```
```{r}
ggsurvplot(
    fit = survfit(Surv(dev.days, pupa.status) ~ host.plant, data = pup.dev), 
    xlab = "Days to Pupation", 
    ylab = "Rate of Pupation",
    fun = "event")
```
```{r}
### non-ggplot plot (lots of easy resources online to make these pretty

cats <- survfit(Surv(dev.days, pupa.status) ~ host.plant, data = pup.dev)
plot(cats,     
    xlab = "days since hatching", 
    ylab = "rate of pupation",
    fun = "event",
    col = c("darkolivegreen", "darkorchid4" ), lwd=3,
    )

legend(0, 1.0, legend=c("Aspen", "Black Cherry"),
       col=c("darkolivegreen", "darkorchid4"),lwd=3, lty=1:1, cex=1.5, title = "Host Plants")

```
Load caterpillar development data using the "curl" command 
```{r}
c <- curl("https://raw.githubusercontent.com/butterfliesrcool/Limenitis-Host-Plant-Gene-Expression/main/caterpillar.development.data.csv")
dev <- read.csv(c, header = TRUE, sep = ",", stringsAsFactors = FALSE)
dev
length(dev$caterpillar.length)
```
```{r}
late.instars <- subset(dev, instar %in% c("4", "5", "pre"), drop = FALSE)
late.instars
length(late.instars$caterpillar.length)
```
```{r}
is.numeric(late.instars$caterpillar.length)
late.instars$caterpillar.length <- as.numeric(late.instars$caterpillar.length)
is.numeric(late.instars$caterpillar.length)
length(late.instars$caterpillar.length)
```
Remove NA data 
```{r}
late.instars$caterpillar.length <- na.omit(late.instars$caterpillar.length)
length(late.instars$caterpillar.length)
```

```{r}
class(late.instars$host.plant)
```

```{r}
late.instars$host.plant <- as.factor(late.instars$host.plant)
class(late.instars$host.plant)
```

```{r}
summary(late.instars$host.plant)
```
```{r}
mean.cat.length <- aggregate(x= late.instars$caterpillar.length, by= list(late.instars$host.plant), FUN=mean)
mean.cat.length
```

```{r}
plot(late.instars$caterpillar.length ~ late.instars$host.plant,
     xlab = "Host Plant",
     ylab = "Late Instar Cateripllar Length",
     col = c("darkolivegreen4", "darkorchid4" ) , lwd=1,
    )
```

```{r}
m <- lm(data = late.instars, caterpillar.length ~ host.plant)
summary(m)
```